---
title: "Mini-Project 2"
author: "Sakaiza RASOLOFOMANANA RAJERY,Phuong Chau, Ann Mudanye"
date: "October 31, 2017"
output: html_document
---


## MINI-PROJECT 2 

**PREDICTION:** 

The amount of money a committee invests on behalf of the candidate in a specific political party affects the number of votes the party earns. 


**DATA SOURCE:** 
 
Throughout this project, we are using files containing committee, candidate and campaign finance data for the 2010-2012 election cycle, that we got from the Federal Election Committee website.  (http://classic.fec.gov/finance/disclosure/ftpdet.shtml#archive_link) 

 

```{r, include=FALSE}
##Loading the date 

load("house_elections.rda")
load("candidates.rda")
load("committees.rda")
load("contributions.rda")
library(tidyverse)
library(dplyr)
library(ggplot2)

```


**FINAL DATA FRAME:**


```{r}
#creating a new committee data frame with the selected columns below
my_committees <- select(committees , cmte_id , cmte_state , cmte_type , cmte_party_affiliation , cmte_filing_freq,org_type , cand_id)

#Rename column cmte_state to state
colnames(my_committees)[colnames(my_committees)=="cmte_state"] <- "state"

#Rename column cmte_party_affiliation to party_affiliation
colnames(my_committees)[colnames(my_committees)=="cmte_party_affiliation"] <- "party_affiliation"

#creating a new contribution data frame with the selected columns below
my_contributions <- contributions %>%
  select(cmte_id, name, state, transaction_amt, transaction_type, cand_id)

#creating a new contribution data frame with the selected columns below
my_candidate <-  candidates %>%
  select(cand_id,cand_name,cand_party_affiliation,cand_election_yr,cand_office_state, cand_name)

# Rename a column in R into state
colnames(my_candidate)[colnames(my_candidate)=="cand_office_state"] <- "state"

#table1 = inner joining my contributions and my comittees by, cmte_id and state
table1 <- inner_join(my_contributions, my_committees, by = c("cmte_id", "state"))

#table2  =  inner joining of my contributions and my candidares by state and cand_id
table2 <- inner_join(my_contributions, my_candidate, by = c("state","cand_id"))

#cleaning table1 by removing cand_id.y column and renaming cand_id.x with cand_id
table1 <- select(table1, -cand_id.y)
colnames(table1)[colnames(table1)=="cand_id.x"] <- "cand_id"

#table3 = inner joining of table1 and 2 by the column below
table3 <- inner_join(table1,table2, by = c("cmte_id", "state", "cand_id", "transaction_amt", "transaction_type", "name"))

#cleaning out table 3 by removing unecessary column
table3 <- select(table3, -party_affiliation, -cmte_type, -cmte_filing_freq, -org_type)


#rename cand_name to candidate name in table 3
colnames(table3)[colnames(table3)=="cand_name"] <- "candidate_name"

#group by candidate name and summarize by transaction amount and candidate id
table3<- table3 %>%
  group_by(candidate_name)%>%
  summarise(total_trans = sum(transaction_amt), candidate_id = cand_id[1])

#create a new data frame my house election by selecting the colomun below
my_house_elections <- select(house_elections,state, candidate_name, party, ge_winner,fec_id)

#rename fec_id with candidate_id in my house election
colnames(my_house_elections)[colnames(my_house_elections)=="fec_id"] <- "candidate_id"

#create a new data frame table4, inner joining my house elections and table3
table4 <- inner_join(table3, my_house_elections, by = "candidate_id") %>%
  select(-candidate_name.y)%>%
  filter(party == "R"|party == "D")

#view the first 6 rows
head(table4)
```
*First six rows of the date frame*


**GRAPH:**

```{r, warning = FALSE}
graph1 <- ggplot(table4, aes(party, log(total_trans), color = ge_winner)) + 
  geom_boxplot() +
  labs(title = "THE DISTRIBUTION OF MONEY TO EACH \nCANDIDATE IN THEIR RESPECTIVE PARTY") +
  theme(plot.title = element_text(hjust = 0, size = 16))+
  ylab(label = "TOTAL AMOUNT RECEIVED") +
  xlab(label = "PARTY")


graph1

```

*N.B:  there are a few outliers with large values on the y-axis causing very small, difficult to represent, interpretations for other data points. Hence the use of a logarithmic scale than a linear scale to avoid this issue.*


In the graph, the x-axis represents two different categorical variables: democracy and republican. We specifically focused the data around these two categorical variables as they are the major parties in the US.
The y-axis indicates the continuous variable: the amount of money in dollars that has been spent on behalf or against a candidate by committees.

The boxplot represents the distribution of people during the election in 2011-2012 as respect to  the amount of money that they received from their committees. The color indicates whether the distribution of a party-- either democracy (with the symbol D) or republican (with the symbol  R) is winning or losing. 

**CONCLUSION:** 

From this graph, we can conclude that it is plausible to win more votes when one candidate has committees spending more money into them. 


